Solar wind prediction for the Solar Probe Plus orbit

### 2017-04-07 paperMVVB/sections/content.tex
implemented cubehelix color scheme for fig. 8 and worked in my offline comments

With his theoretical solar wind model \citet{Parker1958} laid the foundations...\\

Several space missions flew away from Earth's orbit to measure in situ solar wind at different heliocentric distances (e.g. the probes Helios, Voyager, Ulysses...). Voyager~1 is the farthest probe out of the solar system, covering more than 137~au (time? cite?). Ulysses was the single probe orbiting the Sun outside the ecliptic and retrieving solar wind measurements from the poles of the Sun. So far, Helios~2 made the nearest in situ solar wind measurements with \SI{0.29}{\au}, followed closely by Helios~1 with \SI{0.31}{\au}. The new mission Solar~Probe~Plus~(SPP), whose planned launch date is in mid 2018, aims to reach further down into the Sun's atmosphere.

%Solar Probe Plus
The main part of the SPP mission's goals is to gain insight into the coronal heating and solar wind acceleration processes. As these processes occur in the near-Sun region of the corona, SPP will have its closest perihelion at 9.86~solar radii (0.0459~au) \citep{Fox2015}. It will be the first spacecraft flying and measuring in situ solar wind this close to the Sun. Its planned observations range from the perihelion up to \SI{0.25}{\au}, leaving the outer part for data transfer, as \citet{Fox2015} state in their description of SPP's mission design.\\

white light imager WISPR... \citep{Vourlidas2016}\\

There are many studies analyzing the solar wind dependence on solar activity and heliocentric distance with Helios data (\citet{Schwenn1983, Bougeret1984, Schwenn1990}, more cites...). They examine average values, additionally distinguishing between slow and fast wind. In this paper we instead look at the variations of the whole frequency distributions, consider solar activity and extrapolate these to the near-Sun region.\\

@@

In addition, the velocity is the defining parameter of the two types of solar wind. Solar wind mostly consists of and switches between slow and fast wind, whose ratio highly depends on solar activity. These types have different characteristics and their interactions lead to phenomena such as interaction regions. Additionally there also are coronal mass ejections (CMEs) embedded, whose occurrence rate follows solar activity too. Their share varies between almost zero in cycle minimum up to a half in maximum. This study averages over these internal solar wind structures.

@@

Our approach is to get an analytical representation of the frequency distributions' shapes, their solar activity dependence and their solar distance scaling. We get the parameters' frequency distributions and solar activity dependence from near-Earth solar wind and sunspot number (SSN) time series with a duration of almost five solar cycles and their distance dependency from solar wind measurements of more than half a solar cycle covering more than two third of the distance to the Sun (\SIrange{0.29}{0.98}{\au}).%0.68~au\\

@@

%lognormal fitting
%%%%%%%%%%%%%%%%%%
\section{Frequency distributions}
\label{sec:frequency_distribution}
This section looks at the solar wind parameters' frequency distributions which we extract from the in situ OMNI data set. We determine adequate fit function types and evaluate if they are suitable to represent the frequencies' shapes.

\subsection{OMNI frequency data}
\label{sec:omni_frequency_data}
The solar wind parameters are highly variable, due to short-term variations from structures like slow and fast wind streams, interaction regions and CMEs, whose rate and properties depend on solar activity. Hence, for deriving general frequency distributions of the solar wind parameters, averaging over long-time solar wind variations is needed. This requires a distance-independent data set covering multiple solar cycles. The abundance of near-Earth hourly OMNI data is well suited for this task, because it spans almost five solar cycles.

@@

%data source
The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
%OMNIWeb Data Documentation: http://omniweb.gsfc.nasa.gov/html/ow_data.html
% Acknowledgement to the SPDF OMNIWeb database as the source of data used in publications is requested: "The OMNI data were obtained from the GSFC/SPDF OMNIWeb interface at http://omniweb.gsfc.nasa.gov". Further, for recent years when few sources (IMP 8, Wind, ACE, Geotail) contributed to OMNI, it would be appropriate to also cite the PI's who provided the data to OMNI. Copies of preprints or reprints of OMNI-based publications sent to Natalia Papitashvili (address below) would be appreciated for tracking purposes.
% The best citable reference to OMNI data is J.H. King and N.E. Papitashvili, Solar wind spatial scales in and comparisons of hourly Wind and ACE plasma and magnetic field data, J. Geophys. Res., Vol. 110, No. A2, A02209, 10.1029/2004JA010804.
%
%data time range
The hourly data of the whole time range (\textgreater50~years) up to \mbox{2016-12-31} is the basis for the frequency fits. The data starts in \mbox{1963-11-27}, but the temperature data not before \mbox{1965-07-26}. The data coverage of the parameters is between \SIrange{67}{74}{\percent}, which adds up to about 36--40~years in total. This plethora of data is well suited for our task.
%OMNI 1963-01-01 data
%B 1963-11-27 14:30
%V 1963-11-27 12:30
%N 1963-11-27 12:30
%T 1965-07-26 00:30
%1963-01-01~00:00--2016-12-31~23:30 all
%19724 d; 54 y\\
% B: 73.91 \%		39.91 y\\
% V: 73.07 \%		39.46 y\\
% N: 70.29 \%		37.96 y\\
% T: 66.73 \%		36.03 y\\

@@

Obviously all possible values for the four parameters are positive. This hints to the supposition that they are lognormally distributed, as many positive natural quantities conform to a lognormal distribution. This distribution is described by a lognormal function.

@@

	\caption{The resulting fit coefficients from the fitting of the lognormal function (\ref{eq:single_lognormal_fit_function}) to the shape of the solar wind parameters' frequency distributions at \SI{1}{\au}. For the velocity also the fit parameters from the double lognormal function (\ref{eq:double_lognormal_fit_function}) are given, as well as the median and mean values of the resulting velocity fit. The mean absolute errors and sums of absolute residuals are shown as well. The values in brackets are the estimated standard deviation of each fit parameter.}

@@

		\tablefoottext{b}{Velocity median and mean \SI{1}{\au} values for the resulting function. Error estimates derived from the individual fit part errors.}

@@

From visual inspection, the resulting curves match well with the shape of the magnetic field strength, density and temperature distributions (Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}). However, for the velocity the fit function seems insufficient in describing its more complex shape, especially at its peak position and the faster end of the distribution. Its sum of absolute residuals (SAR) between data and fit is almost three times larger than those from the other parameters (Table~\ref{tab:lognormal_fit_parameters}). They can be compared, because the area of probability distributions is unity.
\begin{figure*}
	\includegraphics[width=18cm]{figures/histogram_fits_4_a_zoom_paper_pdfplot.pdf}
	\caption{Frequency distributions of the four solar wind parameters and their lognormal fits. The histograms have bins of 0.5~nT, 10~km/s, 1~cm$^{-3}$ and 10\,000~K and are based on the hourly OMNI data set. The fit's median and mean values are indicated as well. The insets only have zoomed-in frequency axes, their x-axes stay the same.}
	\label{fig:histogram_fits_4_a_zoom_paper_pdfplot}
\end{figure*}
% From the normed stdfit values (Table~\ref{tab:single_lognormal_fit_parameters}) we see that the velocity fit is not that good... \\
% While by visual inspection it is obvious, but also fit stdfit is an indication for a bad fit...\\
% What we see at visual inspection is the area difference between the fit and the data. This area is for the velocity almost three times larger than the other's.\\

@@

The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution gives the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$), which are also listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0
\end{align}
respectively.\\
As anticipated, this more complex fit function is more accurate in describing the velocity's frequency distribution (see Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{Plot of the velocity's frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset only has a zoomed-in frequency axis, its x-axis stays the same.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
In this model the slow and fast part contribute almost equally.\\

In solar cycle maximun (minimum) CMEs contribute with a fraction of \SI{5}{\percent} (half \SI{50}{\percent}) \citep{Richardson2012}.\\

In the following sections we keep using the double lognormal ansatz for the velocity frequency fits...\\

%transition
For the bulk of the solar wind these static lognormal functions describe the parameters' distributions well (the deviation from the fitted functions is less than \SI{7}{\percent}). This is different for the extreme values (which may come from CMEs). The simple lognormal fit models of the magnetic field strength, the velocity and the density underestimate their frequency at the high value tails, whereas the temperature's tail is overestimated (see insets of Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}). The velocity's compositional lognormal fit only slightly overestimates its tail (inset in Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}).\\

@@

\subsection{SSN data}
Solar activity is commonly measured via the sunspot number.\\

activity -- cycle -- forecast -- yearly SSN\\

The international sunspot number is retrieved from the online catalogue\footnote{\url{http://www.sidc.be/silso/}} at the World Data Center -- Sunspot Index and Long-term Solar Observations (WDC-SILSO), Solar Influences Data Analysis Center (SIDC), Royal Observatory of Belgium (ROB).\\
\citet{sidc}\\

why use only yearly SSN data?\\
because OMNI in situ measurements are from Earth orbit (causing variations in solar latitude and distance)\\
seasonal influence contained within yearly frequency distributions\\
yearly OMNI data\\

\subsection{SSN correlation}
We are interested in the correlation of the median, because it defines the position of the lognormal function.\\
yearly parameter medians and yearly SSN (see Fig.~\ref{fig:OMNI_yearly_ssn_correlation_b_plot})\\

Deriving the correlation coefficients for different time lags between solar wind parameters and SSN (see also Fig.~\ref{fig:OMNI_yearly_ssn_correlation_b_plot}).\\

@@

\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_b_plot.pdf}}
	\caption{Plot of the solar wind parameter yearly medians from OMNI data and the yearly SSN from the \citet{sidc} (top). Their correlation coefficients with the yearly SSN for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_b_plot}
\end{figure}

\subsection{SSN fitting}
\label{sec:ssn_fitting}
yearly SSN sw model to sw data fits -> SSN dependency (figs)\\

To shift the distributions with SSN, we fit a linear SSN dependency to the median
\begin{align}
	x_\text{med}(ssn) &= a_\text{med} \cdot ssn + b_\text{med}\,,	\label{eq:median_with_ssn}
\end{align}
using a factor to the SSN $a_\text{med}$ with a baseline $b_\text{med}$. The two values median and mean are sufficient to fully describe lognormal functions, so we relate the mean with a scaling factor to the median to keep the dependency on SSN: (why can we do that? mean proportional to median)
\begin{align}
	x_\text{avg}(ssn) &= (1 + a_\text{avg}) \cdot x_\text{med}(ssn)\,.	\label{eq:mean_with_ssn}
\end{align}
With the implementation of these relations into the lognormal function (\ref{eq:single_lognormal_fit_function}), the new fit function $W'(x,ssn)$ is then fitted to the data. The three resulting fit coefficients ($a_\text{med}, b_\text{med}$ and $a_\text{avg}$) are presented in Table~\ref{tab:ssn_fit_parameters}.\\
\begin{table*}
	\caption{The resulting fit coefficients from the data fitting with SSN. For the velocity the fit parameters from the double lognormal fit and their balancing function are given. The values in brackets are the estimated standard deviation of each fit parameter.}
	\label{tab:ssn_fit_parameters}
	\centering
%	\sisetup{table-figures-integer=1, table-figures-decimal=4, table-figures-exponent=1}
	\begin{tabular}{l@{} c@{}
		S[table-format = 1.3(2)e+1]
		S[table-format = 1.4(2)]
		S[table-format = 1.3(2)e+1]
		S[table-format = +1.3(2)e+1]
		c
		}
		\hline\hline
		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{2}{c}{Median}	&\multicolumn{1}{c}{Mean}	&\multicolumn{2}{c}{Balance}\\
		\cline{3-4}\cline{6-7}
		\multicolumn{2}{l}{}	&\multicolumn{1}{c}{SSN factor $a_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{Baseline $b_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{Scaling factor $a_\text{avg}$}	&\multicolumn{1}{c}{SSN factor $c_a$}	&\multicolumn{1}{c}{Baseline $c_b$}\\
		\hline
		\multicolumn{2}{l}{Magnetic field}	&1.309(19)e-2	&4.285(17)	&8.786(78)e-2	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Velocity}	&2.514(93)e-3	&3.8521(90)	&2.392(26)e-2	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Density}	&3.81(25)e-3	&4.495(26)	&3.050(27)e-1	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Temperature}	&1.974(26)e-2	&5.729(19)	&6.541(28)e-1	&\multicolumn{1}{c}{--}	&--\\
		\hline
		\multirow{2}{*}{Velocity}	&\multicolumn{1}{c}{$W_1$}	&\multicolumn{1}{c}{--}	&3.633(12)	&1.008(37)e-2	&\multirow{2}{*}{\tablenum{-1.799(95)e-3}}	&\multirow{2}{*}{0.638(32)}\\
			&\multicolumn{1}{c}{$W_2$}	&\multicolumn{1}{c}{--}	&4.831(81)	&2.31(20)e-2	&	&\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{Values in their respective units \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}

In comparison with the data the fit models roughly match with the general trends (Fig.~\ref{fig:OMNI_yearly_BVNTSSN_fit_b_plot}).\\
\begin{figure*}
	\includegraphics[width=18cm]{figures/OMNI_yearly_BVNTSSN_fit_b_plot.pdf}
	\caption{Solar wind parameter frequencies, lognormal fit models and the corresponding yearly SSN (grey) over the OMNI time period 1963--2016. The for the model shifted SSN is indicated by a black line.}
	\label{fig:OMNI_yearly_BVNTSSN_fit_b_plot}
\end{figure*}

Again, the velocity gets a special treatment with the double lognormal distribution (see Eq.~\ref{eq:double_lognormal_fit_function}), this time keeping its two components' positions constant and variing instead their balance with the SSN:	(discuss why...)
\begin{align}
	c(ssn) &= c_a \cdot ssn + c_b\,.
\end{align}
The fit result is a model in which three years after solar cycle minimum (SSN of zero) the slow solar wind has a share of almost two-thirds and decreases further with increasing SSN (see Fig.~\ref{fig:Vdbl_SSN_ratio_b_plot}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/Vdbl_SSN_ratio_b_plot.pdf}}
	\caption{Yearly ratio between slow to fast solar wind (threshold \SI{400}{\km\per\s} over SSN (+). Fit model ratio depending on the by 3~years lagged SSN (black line). The relation results from the velocity's double lognormal fit.}
	\label{fig:Vdbl_SSN_ratio_b_plot}
\end{figure}

All fit coefficients are given in Table~\ref{tab:ssn_fit_parameters}).\\

In comparison with the data the fit model roughly matches with the general trend (see Fig.~\ref{fig:OMNI_yearly_VSSN_dblfit_plot}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_VSSN_dblfit_plot.pdf}}
	\caption{Velocity frequency, double lognormal fit model and the corresponding yearly SSN (grey) over the OMNI time period 1963--2016. The for the model shifted SSN is indicated by a black line.}
	\label{fig:OMNI_yearly_VSSN_dblfit_plot}
\end{figure}
3 years after minimum (SSN of zero) the slow wind has its maximal share of 0.638(32). 3 years after maximum (SSN of 200) it has only a share of 0.278(37).\\
...compare with data\\

@@

In this section we use Helios data to obtain exponential fit functions for the heliocentric distance dependency and evaluate the fits' extrapolation behavior in direction to the Sun.

To fit the bulk solar wind distributions' distance dependency we use the frequency fitting method from Sect.~\ref{sec:frequency_distribution} on distance-binned Helios data. This results in solar wind models comprising of exponentially with distance shifted lognormal functions.

@@

As mentioned before, we set the exponents of median and average to be identical. Implementing the exponential distance dependency~(\ref{eq:exponential_function}) into the lognormal function \ref{eq:single_lognormal_fit_function}, we have three fit parameters $a_\text{med}$, $a_\text{avg}$ and the common exponent $b$. (change variable naming...)

@@

Finally we combine the obtained solar activity and distance dependencies for shifting the frequency distributions. The result will be an empirical solar wind model for the inner heliosphere.\\

>>>>

%solar wind models; theoretical and empirical
With his theoretical solar wind model \citet{Parker1958} laid the foundations...\\

%in situ spacecraft
Several space missions flew away from Earth's orbit to measure in situ solar wind at different heliocentric distances (e.g. the probes Helios, Voyager, Ulysses...). Voyager~1 is the farthest probe out of the solar system, now covering more than 137~au (cite?). Ulysses was the single probe orbiting the Sun outside the ecliptic and retrieving solar wind measurements from the poles of the Sun. So far, Helios~2 made the nearest in situ solar wind measurements with \SI{0.29}{\au}, followed closely by Helios~1 with \SI{0.31}{\au}. The new mission Solar~Probe~Plus~(SPP), whose planned launch date is in mid 2018, aims to reach further down into the Sun's atmosphere.

%Solar Probe Plus mission
The main part of the SPP mission's goals is to gain insight into the coronal heating and solar wind acceleration processes. As these processes occur in the near-Sun region of the corona, SPP will have its closest perihelion at 9.86~solar radii (0.0459~au) \citep{Fox2015}. It will be the first spacecraft flying and measuring in situ solar wind this close to the Sun. Its planned observations range from the perihelion up to \SI{0.25}{\au}, leaving the outer part for data transfer, as \citet{Fox2015} state in their description of SPP's mission design.\\

%WISPR
white light imager WISPR... \citep{Vourlidas2016}\\

%difference to other models
There are many studies analyzing the solar wind dependence on solar activity and heliocentric distance with Helios data (\citet{Schwenn1983, Bougeret1984, Schwenn1990}, more cites...). They examine average values, additionally distinguishing between slow and fast wind. In this paper we instead look at the variations of the whole frequency distributions, consider solar activity and extrapolate these to the near-Sun region.\\

@@

In addition, the velocity is the defining parameter of the two types of solar wind. Solar wind mostly consists of and switches between slow and fast wind, whose ratio highly depends on solar activity. These types have different characteristics and their interactions lead to phenomena such as interaction regions. Additionally there also are coronal mass ejections (CMEs) embedded, whose occurrence rate follows solar activity too. Their share varies between almost zero in cycle minimum up to a half in maximum \citep{Richardson2012}. This study averages over these internal solar wind structures.

@@

Our approach is to get an analytical representation of the frequency distributions' shapes, their solar activity dependence and their solar distance scaling. We get the parameters' frequency distributions and solar activity dependence from near-Earth solar wind and sunspot number (SSN) time series with a duration of almost five solar cycles and their distance dependency from solar wind measurements of more than half a solar cycle, covering more than two third of the distance to the Sun (\SIrange{0.29}{0.98}{\au}).%0.68~au\\

@@

%lognormal fitting
%%%%%%%%%%%%%%%%%%
\section{Frequency distributions}
\label{sec:frequency_distribution}
This section looks at the solar wind parameters' frequency distributions which we extract from the in situ OMNI data set. We determine adequate fit function types and evaluate how suitable they are to represent the frequencies' shapes.

\subsection{OMNI frequency data}
\label{sec:omni_frequency_data}
The solar wind parameters are highly variable, due to short-term variations from structures like slow and fast wind streams, interaction regions and CMEs, whose rate and properties depend on solar activity. Hence, for deriving general frequency distributions of the solar wind parameters, averaging over long-term solar wind variations is needed. This requires a distance-independent data set covering multiple solar cycles. The abundance of near-Earth hourly OMNI data is well suited for this task, because it spans almost five solar cycles.

@@

%data source
The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
%OMNIWeb Data Documentation: http://omniweb.gsfc.nasa.gov/html/ow_data.html
% Acknowledgement to the SPDF OMNIWeb database as the source of data used in publications is requested: "The OMNI data were obtained from the GSFC/SPDF OMNIWeb interface at http://omniweb.gsfc.nasa.gov". Further, for recent years when few sources (IMP 8, Wind, ACE, Geotail) contributed to OMNI, it would be appropriate to also cite the PI's who provided the data to OMNI. Copies of preprints or reprints of OMNI-based publications sent to Natalia Papitashvili (address below) would be appreciated for tracking purposes.
% The best citable reference to OMNI data is J.H. King and N.E. Papitashvili, Solar wind spatial scales in and comparisons of hourly Wind and ACE plasma and magnetic field data, J. Geophys. Res., Vol. 110, No. A2, A02209, 10.1029/2004JA010804.
%
%data time range
The hourly data of the whole time range up to \mbox{2016-12-31} is the basis for the frequency fits. The data starts in \mbox{1963-11-27}, but the temperature data not before \mbox{1965-07-26}. The data coverage of the parameters is between \SIrange{67}{74}{\percent}, which adds up to about 36--40~years in total. This plethora of data is well suited for our task.
%OMNI 1963-01-01 data
%B 1963-11-27 14:30
%V 1963-11-27 12:30
%N 1963-11-27 12:30
%T 1965-07-26 00:30
%1963-01-01~00:00--2016-12-31~23:30 all
%19724 d; 54 y\\
% B: 73.91 \%		39.91 y\\
% V: 73.07 \%		39.46 y\\
% N: 70.29 \%		37.96 y\\
% T: 66.73 \%		36.03 y\\

@@

Obviously all possible values for the four parameters are positive. This hints to the supposition that they are lognormally distributed, as many positive natural quantities conform to lognormal distributions. Its probability density function is described by a lognormal function.

@@

	\caption{Resulting fit coefficients from the fitting of the lognormal function (\ref{eq:single_lognormal_fit_function}) to the shape of the solar wind parameters' frequency distributions at \SI{1}{\au}. For the velocity also the fit parameters from the double lognormal function (\ref{eq:double_lognormal_fit_function}) are given, as well as the median and mean values of the resulting velocity fit. The mean absolute errors and sums of absolute residuals are shown as well. The values in brackets are the estimated standard deviation of each fit parameter.}

@@

		\tablefoottext{b}{Error estimates derived from the individual fit part errors.}

@@

From visual inspection, the resulting curves match well with the shape of the magnetic field strength, density and temperature distributions (Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}). However, for the velocity the fit function seems insufficient in describing its more complex shape, especially at its peak position and the faster end of the distribution. Its sum of absolute residuals (SAR) between data and fit is almost three times larger than those from the other parameters (Table~\ref{tab:lognormal_fit_parameters}). They can be compared, because the area of probability density functions is unity.
\begin{figure*}
	\includegraphics[width=18cm]{figures/histogram_fits_4_a_zoom_paper_pdfplot.pdf}
	\caption{Frequency distributions of the four solar wind parameters and their lognormal fits. The histograms have bins of 0.5~nT, 10~km/s, 1~cm$^{-3}$ and 10\,000~K and are based on the hourly OMNI data set. The fit's median and mean values are indicated as well. The insets only have zoomed-in frequency axes, their x-axes stay the same.}
	\label{fig:histogram_fits_4_a_zoom_paper_pdfplot}
\end{figure*}
% From the normed stdfit values (Table~\ref{tab:single_lognormal_fit_parameters}) we see that the velocity fit is not that good... \\
% While by visual inspection it is obvious, but also fit stdfit is an indication for a bad fit...\\
% What we see at visual inspection is the area difference between the fit and the data. This area is for the velocity almost three times larger than the other's.\\

@@

The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution gives the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$), which are also listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0
\end{align}
respectively.\\
As anticipated, this more complex fit function is more accurate in describing the velocity's frequency distribution (see Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{Plot of the velocity's frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset only has a zoomed-in frequency axis, its x-axis stays the same.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
That is why we keep using the double lognormal ansatz for the velocity frequency fits in the following sections.\\
In this static model the slow and fast part contribute almost equally ($c \approx 0.5$), which of course is only valid for this kind of long-term average. At different times in a solar cycle their contributions vary strongly.\\

%transition
For the bulk of the solar wind these static lognormal functions describe the parameters' distributions well. This is different for the extreme values (which may also stem from CMEs). The simple lognormal fit models of the magnetic field strength, the velocity and the density underestimate their frequency at the high value tails, whereas the temperature's tail is overestimated (see insets of Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}). The velocity's compositional lognormal fit only slightly overestimates its tail (inset in Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}).\\

@@

\subsection{SSN data}
Solar activity is commonly measured via the sunspot number. We want to correlate OMNI in situ measurements with the SSN, but OMNI data are from Earth orbit, causing variations in solar latitude and distance. To dodge these variations we use yearly OMNI and SSN data.\\

The international sunspot number (\citeyear{sidc}) is retrieved from the online catalogue\footnote{\url{http://www.sidc.be/silso/}} at the World Data Center -- Sunspot Index and Long-term Solar Observations (WDC-SILSO), Solar Influences Data Analysis Center (SIDC), Royal Observatory of Belgium (ROB).\\

seasonal influence contained within yearly frequency distributions\\

SIDC and SWPC provide limited SSN forecast (a few months)\\

\subsection{SSN correlation}
Our current interest lies in the correlation of the SSN to the solar wind median value, because it defines the position of the lognormal function. The yearly OMNI parameter medians and the yearly SSN are plotted in Fig.~\ref{fig:OMNI_yearly_ssn_correlation_b_plot}.\\

solar activity = SSN $\neq$ solar cycle\\
The solar wind velocity (and its close friends N \& T) is known to depend on the state of the solar cycle (cite?), which is why it follows the SSN only indirectly (with time lag). Thus we derive the correlation coefficients for different time lags between solar wind parameters and SSN (see Fig.~\ref{fig:OMNI_yearly_ssn_correlation_b_plot}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_b_plot.pdf}}
	\caption{Plot of the solar wind parameter yearly medians from OMNI data and the yearly SSN from the \citet{sidc} with cycle number (top). Their correlation coefficients with the yearly SSN for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_b_plot}
\end{figure}

@@

yearly SSN sw model to sw data fits -> SSN dependency (figs)\\

\subsection{SSN fitting}
\label{sec:ssn_fitting}
To be able to shift the distributions with SSN, add fit a linear SSN dependency to the median
\begin{align}
	x_\text{med}(ssn) &= a_\text{med} \cdot ssn + b_\text{med}\,,	\label{eq:median_with_ssn}
\end{align}
using a factor to the SSN $a_\text{med}$ with a baseline $b_\text{med}$. We relate the mean with a scaling factor to the median to keep the new SSN dependency:
\begin{align}
	x_\text{avg}(ssn) &= (1 + a_\text{avg}) \cdot x_\text{med}(ssn)\,.	\label{eq:mean_with_ssn}
\end{align}
With the implementation of these relations into the lognormal function (\ref{eq:single_lognormal_fit_function}), the new dynamic fit function $W'(x,ssn)$ is then fitted to the yearly data. The three resulting fit coefficients ($a_\text{med}, b_\text{med}$ and $a_\text{avg}$) are presented in Table~\ref{tab:ssn_fit_parameters}.\\
\begin{table*}
	\caption{Resulting fit coefficients from the data fitting with SSN. For the velocity the fit parameters from the double lognormal fit and their balancing function are given. The values in brackets are the estimated standard deviation of each fit parameter.}
	\label{tab:ssn_fit_parameters}
	\centering
%	\sisetup{table-figures-integer=1, table-figures-decimal=4, table-figures-exponent=1}
	\begin{tabular}{l@{} c@{}
		S[table-format = 1.3(2)e+1]
		S[table-format = 1.4(2)]
		S[table-format = 1.3(2)e+1]
		S[table-format = +1.3(2)e+1]
		c
		}
		\hline\hline
		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{2}{c}{Median}	&\multicolumn{1}{c}{Mean}	&\multicolumn{2}{c}{Balance}\\
		\cline{3-4}\cline{6-7}
		\multicolumn{2}{l}{}	&\multicolumn{1}{c}{SSN factor $a_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{Baseline $b_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{Scaling factor $a_\text{avg}$}	&\multicolumn{1}{c}{SSN factor $c_a$}	&\multicolumn{1}{c}{Baseline $c_b$}\\
		\hline
		\multicolumn{2}{l}{Magnetic field}	&1.309(19)e-2	&4.285(17)	&8.786(78)e-2	&\multicolumn{1}{c}{--}	&--\\
%		\multicolumn{2}{l}{Velocity}	&2.514(93)e-3	&3.8521(90)	&2.392(26)e-2	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Density}	&3.81(25)e-3	&4.495(26)	&3.050(27)e-1	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Temperature}	&1.974(26)e-2	&5.729(19)	&6.541(28)e-1	&\multicolumn{1}{c}{--}	&--\\
		\hline
		\multirow{2}{*}{Velocity}	&\multicolumn{1}{c}{$W_1$}	&\multicolumn{1}{c}{--}	&3.633(12)	&1.008(37)e-2	&\multirow{2}{*}{\tablenum{-1.799(95)e-3}}	&\multirow{2}{*}{0.638(32)}\\
			&\multicolumn{1}{c}{$W_2$}	&\multicolumn{1}{c}{--}	&4.831(81)	&2.31(20)e-2	&	&\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{Values in their respective units \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}

In comparison with the data the fit models seem to match with the general trends (Fig.~\ref{fig:OMNI_yearly_BVdblNTSSN_fit_d_plot}).\\
\begin{figure*}
	\includegraphics[width=18cm]{figures/OMNI_yearly_BVdblNTSSN_fit_d_plot.pdf}
	\caption{Solar wind parameter data frequencies, lognormal fit models with their median values (white) and the corresponding yearly SSN (grey) over the OMNI time period 1963--2016. The for the models shifted SSN is indicated by a black line. The constant medians of both velocity's lognormal parts are marked with white dots.}
	\label{fig:OMNI_yearly_BVdblNTSSN_fit_d_plot}
\end{figure*}

Again, the velocity gets a special treatment with the double lognormal distribution (\ref{eq:double_lognormal_fit_function}). It is known that slow and fast solar wind stream occurrence rates follow the solar cycle, but their magnitudes stay fairly stable (cite?). This time we keep the two velocity components' positions constant and vary instead their balance with the SSN:
\begin{align}
	c(ssn) &= c_a \cdot ssn + c_b\,.
\end{align}
The fit result (see Table~\ref{tab:ssn_fit_parameters}) is a model in which three years after solar cycle minimum (SSN of zero) the slow solar wind has a share of almost two-thirds and decreases further with increasing SSN (see Fig.~\ref{fig:Vdbl_SSN_ratio_b_plot}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/Vdbl_SSN_ratio_b_plot.pdf}}
	\caption{Yearly ratio between slow to fast solar wind (threshold \SI{400}{\km\per\s} over SSN (+). Fit model ratio depending on the by 3~years lagged SSN (black line). The relation results from the velocity's double lognormal fit.}
	\label{fig:Vdbl_SSN_ratio_b_plot}
\end{figure}
% 3 years after minimum (SSN of zero) the slow wind has its maximal share of 0.638(32). 3 years after maximum (SSN of 200) it has only a share of 0.278(37).\\
...compare with data\\
=> there is no specific velocity threshold between slow and fast solar wind types, the velocity ranges of both types overlap.\\

In comparison with the data the fit model roughly matches with the general trend (see Fig.~\ref{fig:OMNI_yearly_BVdblNTSSN_fit_d_plot}).\\

@@

In this section we use Helios data to obtain exponential fit functions for the heliocentric distance dependency and evaluate the fits' extrapolation behavior in direction to the Sun. To fit the bulk solar wind distributions' distance dependency we use the frequency fitting method from Sect.~\ref{sec:frequency_distribution} on distance-binned Helios data. This results in models comprising of exponentially with distance shifted lognormal functions.

@@

As mentioned before, we set the exponents of median and average to be identical. Implementing the exponential distance dependency~(\ref{eq:exponential_function}) into the lognormal function (\ref{eq:single_lognormal_fit_function}), we get three fit parameters ($a_\text{med}$, $a_\text{avg}$ and the common exponent $b$).

@@

Finally, we combine the obtained solar activity and distance dependencies for shifting the frequency distributions. The result will be an empirical solar wind model for the inner heliosphere.\\
